\propertitle

### 2019-07-20 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-20T12:48:31+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain areas \mynote{be
  more specific}. For many animal species, the neurons that are recorded
within the area are not -- and at present cannot be -- specifically chosen
from among the rest, owing to several limiting factors; for example,
limitations in how precisely electrodes are inserted or neurons are
targeted by viruses. In fact, the set of recorded neurons may even change
during very long recordings or across experimental sessions. We assume that
there's an area, comprising a network of $N$ neurons, from which other sets
of neurons could have been or will be recorded***have the same probability of being recorded, even in
other experimental sessions, as the set of $n$ neurons that was actually
recorded in this session. This is our \enquote{full network}.\mynote{how
  about calling it \enquote{the pool}?} The normalized total activity of
these $N$ neurons at $t$ is $\yA_{t}$. The relative frequencies of the
various activity levels during the recording were $(\yFF_{\yA}) \defs \yF$.
We don't know the values $\yA_{t}$ at each $t$, or the frequency
distribution $\yF$. We only know for certain that
$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
for obvious reasons. For the time being we assume that we know $N$; in
\sect*** we discuss the consequences of our lack of precise knowledge about
this number.

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-20T18:44:47+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain areas \mynote{be
  more specific}. For many animal species, the neurons that are recorded
within the area are not -- and at present cannot be -- specifically chosen
from among the rest, owing to several limiting factors; for example,
limitations in how precisely electrodes are inserted or neurons are
targeted by viruses. In fact, the set of recorded neurons may even change
during very long recordings or across experimental sessions. We assume that
there's an area, comprising a network of $N$ neurons, for which we equally
believe that any other sample of size $n$ could have been recorded instead
of the sample of $n$ neurons that was actually recorded. This is our
\enquote{full network}.\mynote{how about calling it \enquote{the pool}?}
The normalized total activity of these $N$ neurons at $t$ is $\yA_{t}$. The
relative frequencies of the various activity levels during the recording
were $(\yFF_{\yA}) \defs \yF$. We don't know the values $\yA_{t}$ at each
$t$, or the frequency distribution $\yF$. We only know for certain that
$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
for obvious reasons. For the time being we assume that we know $N$; in
\sect*** we discuss the consequences of our lack of precise knowledge about
this number.

@@

\section{The full-network size $N$}
\label{sec:N}

The calculations and conclusions presented in the preceding sections depend
on the size of the full network, $N$. The full network can't be the whole
brain, of course; that would be silly. How large can $N$ be?

The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
our method is based, and which asserts the equality of the factorial
moments of the full network and its sample, (or, equivalently, asserting
that the power moments for the one are linear functions of the power
moments for the other). This formula is only valid if our beliefs about the
activity of the sample  $\pxx(\ya)$ and that of the full network
$\pxx(\yA)$ are related by a hypergeometric distribution:
\begin{equation}
  \label{eq:hypergeometric_relation}
    \pxx(\ya)=
  \binom{n}{n\ya}\;\binom{\yNv-n}{\yNv \yA-n\ya}\;\binom{\yNv}{\yNv \yA}^{-1}
\;\pxx(\yA).
\end{equation}
This is a relation of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, $N$ must be such that we equally believe other samples of
size $n$ could have been recorded by the electrodes or the probe used. This
is indeed what we stated in \sect~\ref{sec:model}.

This requirement delimits an area around the probe. Its number of neurons
depends on the animal species and on the brain region from which the
recording was made, as the neuron density can be very different. It's
practically impossible to specify the exact number of neurons, but its
order of magnitude is enough for making qualitative inferences. As we'll
discuss in the next section, even if the exact number were known the
maximum-entropy distribution ought still to be interpreted qualitatively.

It would be possible to derive a distribution $\pxx(A)$ for the full
network based on an unknown $N$, by  expressing our belief distribution
$\pxx(N)$ about $N$ and using this to marginalize $N$ out:
\begin{equation}
  \label{eq:marginalize_N}
  \pxx(A \| \text{$N$ \small unknown}) = \sum_{N} \px(\yA \| N)\;\pxx(N).
\end{equation}
But this would be overkill, owing to the qualitative character of the
maximum-entropy distribution constructed with exact $N$.
